### 2020-06-25 large_experiment_processing.Rmd
feat: add supplementary data showcasing MSnbase on a large data set

---
title: "Supplementary data for: \\textit{`MSnbase`, efficient and elegant R-based processing and visualisation of raw mass spectrometry data}"
author: "Laurent Gatto, Sebastian Gibb and Johannes Rainer"
output:
    BiocStyle::pdf_document
---

```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

This document describes handling of mass spectrometry data from large
experiments using the `MSnbase` package and more specifically its *on-disk*
backend. For demonstration purposes, the
[MassIVE](https://massive.ucsd.edu/ProteoSAFe/static/massive.jsp) data set
[MSV000080030](https://massive.ucsd.edu/ProteoSAFe/dataset.jsp?task=5e7034cc98c54a47b803b144bff6a296)
is used. This consists of over 1,000 mzXML files from swab-samples collected
from hands and various personal objects of 80 volunteers.

# Data handling and analysis with `MSnbase`

In this section we demonstrate data handling and access by `MSnbase` on a large
experiment consisting of more than 1,000 data files.

To reproduce the analysis in this document, download the *MSV000080030* folder
from
[ftp://massive.ucsd.edu/MSV000080030/](ftp://massive.ucsd.edu/MSV000080030/) and
place it into the same folder than this document.

Below we load the required libraries and define the files to be analyzed.

```{r}
library(MSnbase)
library(magrittr)
library(pryr)

fls <- dir("MSV000080030/ccms_peak/Forensic_study_80_volunteers/",
           pattern = "mzXML", full.names = TRUE, recursive = TRUE)
```

The data set consists of in total `r length(fls)` mzXML files. We next load the
data using the two different `MSnbase` backends `"inMemory`" and `"onDisk"`. For
the in-memory backend, due to the larger memory requirements, we import the data
only from a subset of the files.

```{r import-inmem, eval = !file.exists("ms_mem_hand.RData")}
ms_mem <- readMSData(fls[grep("Hand", fls)], mode = "inMemory")
```

```{r import-inmem-save, echo = FALSE}
if (!file.exists("ms_mem_hand.RData")) {
    save(ms_mem, file = "ms_mem_hand.RData")
} else {
    load("ms_mem_hand.RData")
}
```

Next we load data from all mzXML files as an on-disk `MSnExp` object.

```{r import-ondisk, eval = !file.exists("ms_dsk_all.RData")}
ms_dsk <- readMSData(fls, mode = "onDisk")
```

```{r import-ondisk-save, echo = FALSE}
if (!file.exists("ms_dsk_all.RData")) {
    save(ms_dsk, file = "ms_dsk_all.RData")
} else {
    load("ms_dsk_all.RData")
}
```

Below we count the number of spectra per MS level of the whole experiment.

```{r ondisk-spectra-table}
table(msLevel(ms_dsk))
```

Note that the in-memory `MSnExp` object contains only MS2 spectra (in total 
`r length(ms_mem)`) from a subset of data files, still, data import was much 
slower (over ~ 12 hours for the in-memory backend while creating the on-disk 
object from the full data data set took ~ 3 hours).

Next we subset the on-disk object to contain the same set of spectra than the
in-memory `MSnExp` and compare their memory footprint.

```{r memory-footprint}
ms_dsk_hands <- ms_dsk %>%
    filterFile(grep("Hand", fls)) %>%
    filterMsLevel(2L)

object_size(ms_mem)
object_size(ms_dsk_hands)
```

Since the on-disk object stores only spectra metadata in memory it occupies also
much less system memory. As a comparison, the on-disk `MSnExp` for the full
experiment was still much smaller than the in-memory object:

```{r memory-footprint-all}
object_size(ms_dsk)
```

## Performance of the *on-disk* backend on large scale data sets

To demonstrate `MSnbase`'s efficiency in processing large scale experiments we
perform some standard subsetting, data access and manipulation operations.

```{r parallel-setup, echo = FALSE}
library(BiocParallel)
register(SerialParam())
```

We first compare the performance of the on-disk and in-memory backend on
accessing m/z values with the `mz` function on a set of 100 randomly selected
spectra. The performance is assessed with the `microbenchmark` function.

```{r random-mz}
set.seed(123)
idx <- sample(seq_along(ms_mem), 100)

library(microbenchmark)
microbenchmark(mz(ms_mem[idx]),
               mz(ms_dsk_hands[idx]),
               times = 5)
```

Thus, for this combined subsetting and data access operation the on-disk backend
performed better than the in-memory `MSnExp`, while even requiring much less
memory.

Next we extract all MS2 spectra with a retention time between 50 and 60 seconds
and a precursor m/z of 108.5362 (+/- 5ppm). This subsetting operation is
performed on the on-disk `MSnExp` object representing the full experiment with
the `r length(fileNames(ms_dsk))` data files/samples. To assess the performance
of the following operations we use `system.time` calls that record elapsed time
in seconds.

```{r precursor-selection}
system.time(
    ms_sub <- ms_dsk %>%
        filterMsLevel(2L) %>%
        filterRt(c(50, 60)) %>%
        filterPrecursorMz(mz = 108.5362, ppm = 5)
)["elapsed"]
```

In total `length(ms_sub)` spectra were selected from in total 
`r length(unique(fromFile(ms_sub)))` data files/samples. The plot below shows 
the data for the first spectrum.

```{r precursor-selection-plot-1}
system.time(
    plot(ms_sub[[1]])
)["elapsed"]
```

Since there seems to be quite some background noise in the MS2 spectrum we next
remove peaks with an intensity below 50 by first replacing their intensities
with 0 (with the `removePeaks` call) and subsequently removing all 0-intensity
peaks from each spectrum with the `clean` call. In addition we *normalize* each
spectrum by dividing the maximum intensity per spectrum from the spectrum's
intensities.

```{r normalize}
system.time(
    ms_sub <- ms_sub %>%
        removePeaks(t = 50) %>%
        clean(all = TRUE) %>%
        normalize(method = "max")
)["elapsed"]
```

The result on the first spectrum is shown below.

```{r precursor-selection-plot-normalized}
system.time(
    plot(ms_sub[[1]])
)["elapsed"]
```

Note that any of the data manipulations above are not directly applied to the
data but *cached* in the object's internal *lazy processing queue* (explaining
the very short running time of the normalization call). The operations are only
effectively applied to the data when m/z or intensity values are extracted from
the object as it is the case in the `plot` call above.

For additional workflows employing `MSnbase` see also
[metabolomics2018](https://jorainer.github.io/metabolomics2018/xcms-preprocessing.html)[^1]
that explains filtering, plotting and centroiding of profile-mode MS data with
`MSnbase` and subsequent pre-processing of the (label free/untargeted) LC-MS
data with the `xcms` package (that builds upon `MSnbase` for MS data
representation and access).

## System information

The present analysis was run on a MacBook Pro 16,1 with 2.3 GHz 8-Core Intel
Core i9 CPU and 64 GB 2667 MHz DDR4 memory running macOS version 10.15.5. The R
version and the version of the used packages are listed below.

```{r session-info}
sessionInfo()
```

[^1]: https://github.com/jorainer/metabolomics2018